\chapter{Proof Assistants}\label{app:ProofAssist}

Initially, my goal was to learn how to use proof assistants to assert properties about Proteus.
I began to learn how to write basic and simple proofs using Coq.
Coq was powerful enough to show properties of systems that were publishable.
Whilst learning how to use Coq, I was learning to use TLA+ which proved equally as powerful, but not suited for my use case.
TLA+ is used for simultaneous computing as well as state machines which seemed very appealing upon initial inspection, since Proteus uses state machines as well.
However, this didn't extend well into demonstrating parts of Proteus that were not related to state machines.
I then considered using a different proof assistant like Dafny, Lean, or Twelf to see which would be better suited for proving something about Proteus.
I continued learning how to use the proof assistants and aimed to apply them to whatever proof I would eventually create about Proteus.

The big idea hit me one day; To show that Proteus was Turing Complete.
It was one of the biggest things to show in theory and was able to utilize all my knowledge and experience from this discipline.
There was only 1 problem, I couldn't use any proof assistant.
This is because fundamentally, it is impossible to show Turing Completeness using a proof assistant.

Recall Turing Completeness means that the shown system is equivalently as powerful as a TM, see section \ref{sec:TC}.
Also, recall the Church-Turing Thesis, Theorem \ref{thm:CTT}, which means that the definition of a TM is to be able calculate any calculable function.
Additionally, recall the proof of showing a system is TC in section \ref{subsubsec:NotableEgsFormalLang}.
The purpose of a proof assistant is to assert that the given proof is valid, step by step until the goal is reached.
Therefore for all inputs for any proof assistant, they are decidable (assuming they are well-written).
Thus, they are not capable of satisfying the Church-Turing Thesis, Theorem \ref{thm:CTT}, and also cannot be as powerful as a TM.
As such, it is impossible to use a proof assistant to show Turing Completeness.
This is when I abandoned using a proof assistant and moved forward in favor of proving Proteus TC.