\documentclass[12pt]{report} % use report for thesis formatting and setup with 12 pt font
\usepackage[american]{babel} % use standard american english for hyphenation rules, date format, etc.
\usepackage{mathptmx} % Times New Roman
\usepackage{graphicx} % for importing images
\usepackage{cite} % referencing citations in bib
\usepackage[doublespacing]{setspace} %\doublespacing % set double spaced for entire doc
\usepackage[margin=1in]{geometry} % 1 inch margins
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} % used for website linkes to be displayed correctly
\usepackage{etoolbox} % used for changing vertical space before chapter headers to 0
\usepackage{titlesec} % customize title of TOC, and modify sections/subsections/etc.
\usepackage{sectsty} % used for centering the chapters and text for chapter
\usepackage{tocloft} % adjust spacing in TOC. adjust chapter and titles to be not bold in TOC
\usepackage{amsmath} % insert text inside of math sections
\usepackage{amssymb} % math symbols like Naturals, Integers, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT SETUP

% displays subsubsection to TOC and doc
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

% create the subsubsection
\titleformat{\subsubsection}[runin]
{\normalfont\normalsize\bfseries}
{\thesubsubsection}
{0.6em}
{}

% change bibliography to references
\addto\captionsamerican{\renewcommand{\bibname}{References}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FORMATTING FOR EACH SECTION

% remove vertical space before & after the TOC title
\setlength{\cftbeforetoctitleskip}{0em}
\setlength{\cftaftertoctitleskip}{0em}

% redefine makechapterhead to remove spacing and apply font styles
\makeatletter
\def\@makechapterhead#1{
  {\parindent \z@ \raggedright \normalfont
    \ifnum \c@secnumdepth >\m@ne
        \centering \bfseries \normalsize \@chapapp\space \thechapter % center and normal size font with bold typesetting
        \par\nobreak
        \vskip 20\p@
    \fi
    \interlinepenalty\@M
    \normalsize #1\par\nobreak % normal size font
    \vskip 40\p@
  }}
\def\@makeschapterhead#1{
  {\parindent \z@ \raggedright
    \centering % centered text
    \normalfont % normal sized font
    \interlinepenalty\@M
    \normalsize  #1\par\nobreak % normal sized font
    \vskip 40\p@
  }}
\makeatother

% set each section/subsection/subsubsection to normal sized font
\sectionfont{\normalsize}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalsize}

% change TOC name, center horizontally, and make non bold
\addto\captionsamerican{
    \renewcommand{\contentsname}{\hfill Table of Contents \hfill}
    \renewcommand{\cfttoctitlefont}{\normalsize}
    \renewcommand{\cftaftertoctitle}{\hfill}
}

% change chapters to be non-bold in TOC
\renewcommand{\cftchapfont}{\normalfont}
\renewcommand{\cftchappagefont}{\normalfont}

% setup for definitions, theorems, etc.
\newtheorem{thm}{Theorem}
\newtheorem{definition}{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN DOCUMENT

\begin{document}
    % title page & begin roman numeral numbering
    \pagenumbering{roman}
    \begin{titlepage}
        \centering
        CALIFORNIA STATE UNIVERSITY, NORTHRIDGE\\
        [1.5in]
        %Title
        Survey of Imperative Style Turing Complete proof techniques\\
        and an application to prove Proteus Turing Complete\\
        [1.5in]
        A thesis submitted in partial fulfillment of the requirements\\
        for the degree of Master of Science in\\
        Computer Science\\
        [0.5in]
        By\\
        [0.25in]
        Isaiah Martinez\\
        \vspace{\fill}
        December 2024
    \end{titlepage}

    \newpage
    % unsigned Signature Page
    \setcounter{page}{2} % set page counter in roman numerals to ii

    \chapter*{}\label{chapter:SignPage}
    \addcontentsline{toc}{chapter}{Signature Page}
        
        
        %this horizontal spacing keeps changing so just make sure it looks good in the pdf
        %such that it matches the vertical line to be above the 'K' in Kyle
        \hspace{0.9in}
        The thesis of Isaiah Martinez is approved:
        \vspace{2in}

        \begin{center}
            \begin{tabular}{p{3in} p{1in}}
                \rule{3in}{0.4pt} & \rule{1in}{0.4pt}\\
                Maryam Jalali, PhD. & Date \\
            \end{tabular}
        \end{center}

        \begin{center}
            \begin{tabular}{p{3in} p{1in}}
                \rule{3in}{0.4pt} & \rule{1in}{0.4pt}\\
                John Noga, PhD. & Date \\
            \end{tabular}
        \end{center}

        \begin{center}
            \begin{tabular}{p{3in} p{1in}}
                \rule{3in}{0.4pt} & \rule{1in}{0.4pt}\\
                Kyle Dewey, PhD., Chair & Date \\
            \end{tabular}
        \end{center}

        \vspace{\fill}

        \begin{center}
            California State University, Northridge
        \end{center}

    \newpage

    % acknowledgements
    % \chapter*{\normalfont Acknowledgements}\label{chapter:Acknowledgements}
    % \addcontentsline{toc}{chapter}{Acknowledgements}
    % \vspace{3in}
    %     \begin{center}
    %         Lorem ipsum dolor sit amet, consectetur adipiscing elit.
    %         Vivamus lacinia odio vitae vestibulum vestibulum.
    %         Cras venenatis euismod malesuada.
    %         Maecenas vehicula felis quis eros auctor, sed efficitur erat suscipit.
    %         Curabitur vel lacus velit. Proin a lacus at arcu porttitor vehicula.
    %         Mauris non velit vel lectus tincidunt ullamcorper at id risus.
    %         Sed convallis sollicitudin purus a scelerisque.
    %         Phasellus faucibus purus at magna tempus, sit amet aliquet nulla cursus.
    %     \end{center}
    

    \newpage

    \tableofcontents

    \newpage

    % List of figures/tables/illustrations
    % Delete any of the following 3 "List of __" that have < 4 items or are not scattered throughout the text

    % \chapter*{\normalfont List of Figures}\label{chapter:ListFigures}
    % \addcontentsline{toc}{chapter}{List of Figures}
    % \begin{center}
    %     This list must reference the figure, page it appears, and subject matter.
    % \end{center}
    
    % \newpage

    % \chapter*{\normalfont List of Illustrations}\label{chapter:ListIllustrations}
    % \addcontentsline{toc}{chapter}{List of Illustrations}
    % \begin{center}
    %     This list must reference the illustration, page it appears, and subject matter.
    % \end{center}
    
    \newpage

    \input{Abstract}

    \newpage

    \pagenumbering{arabic} % start normal page counting
    \input{Intro}
    \input{DiffTech}
    \input{ProteusisTC}
    \input{Conclusion}

    \newpage

    % bibliography
    
    %alphabetical order
    %retrieval date is only for non-journal instances where material might change at a later date

    \begin{singlespace}

        \begin{thebibliography}{}

            \bibitem{NDTMeqDTM} Q. Gao and X. Xu, “The Analysis and Research on Computational Complexity,” pp. 3467–3472.

            \bibitem{GandyPaper} R. Gandy, “Church’s Thesis and Principles for Mechanisms,” The Kleene Symposium, pp. 123–148, Jun. 1980.

            \bibitem{ProteusRunTime} B. McClelland, “Adding Runtime Verification to the Proteus Language,” CSUN, May 2021.

            \bibitem{IntroFormLangAuto} P. Linz, An Introduction to Formal Languages and Automata. Jones \& Bartlett Learning, 2016.

            \bibitem{JavaGenericsTC} R. Grigore, “Java generics are turing complete,” ACM SIGPLAN Notices, vol. 52, no. 1, pp. 73–85, Jan. 2017, doi: https://doi.org/10.1145/3093333.3009871.

            \bibitem{MtGTC} A. Churchill, S. Biderman, and A. Herrick, “Magic: The Gathering is Turing Complete.”

            \bibitem{CivTC} A. de Wynter, “Turing Completeness and Sid Meier’s Civilization,” IEEE Transactions on Games, vol. 15, no. 2, pp.292-299, June 2023.

            \bibitem{GDLTC} A. Saffidine, “The Game Description Language Is Turing Complete,” IEEE Transactions on Computational Intelligence and AI in Games, vol. 6, no. 4, pp. 320–324, Dec. 2014, doi: https://doi.org/10.1109/tciaig.2014.2354417.

            \bibitem{CGoLTM} P. Rendell, "A Universal Turing Machine in Conway's Game of Life," 2011 International Conference on High Performance Computing \& Simulation, Istanbul, Turkey, 2011, pp. 764-772, doi: 10.1109/HPCSim.2011.5999906.
        
            \bibitem{TTTTM} S. S. T. Gontumukkala, Y. S. V. Godavarthi, B. R. R. T. Gonugunta and S. M., "Implementation of Tic Tac Toe Game using Multi-Tape Turing Machine," 2022 International Conference on Computational Intelligence and Sustainable Engineering Solutions (CISES), Greater Noida, India, 2022, pp. 381-386, doi: 10.1109/CISES54857.2022.9844404.

            \bibitem{ShallityAutomataThy} Jeffrey Outlaw Shallit, A second course in formal languages and automata theory. Cambridge ; New York: Cambridge University Press, 2009.

            \bibitem{JFLAPGrading} M. Biçer, F. Albayrak and U. Orhan, "Automatic Automata Grading System Using JFLAP," 2023 Innovations in Intelligent Systems and Applications Conference (ASYU), Sivas, Turkiye, 2023, pp. 1-4, doi: 10.1109/ASYU58738.2023.10296744.

            \bibitem{TuBB} E. Luce and S. H. Rodger, "A visual programming environment for Turing machines," Proceedings 1993 IEEE Symposium on Visual Languages, Bergen, Norway, 1993, pp. 231-236, doi: 10.1109/VL.1993.269602.
            
            \bibitem{LambdaCalcKleene} Dezani‐Ciancaglini Mariangiola and J. R. Hindley, “Lambda-Calculus,” Wiley Encyclopedia of Computer Science and Engineering, pp. 1–8, Sep. 2008, doi: https://doi.org/10.1002/9780470050118.ecse212.

            \bibitem{LambdaCalcRG} M. Dezani-Ciancaglini and J. R. Hindley, “Lambda-Calculus,” Nov. 2007, Available: \href{https://www.researchgate.net/publication/228107078_Lambda-Calculus}{https://www.researchgate.net/publication/228107078\_Lambda-Calculus}

            \bibitem{TutLambdaCalc} R. Rojas, “A Tutorial Introduction to the Lambda Calculus,” 2015, Available: \href{https://arxiv.org/pdf/1503.09060}{https://arxiv.org/pdf/1503.09060}

            \bibitem{AckermannPR} CWoo, “Ackermann function is not primitive recursive,” Mar. 2013, Available: \href{https://www.cs.tau.ac.il//~nachumd/term/42019.pdf}{https://www.cs.tau.ac.il//~nachumd/term/42019.pdf}
        
            \bibitem{RecursiveFuncs} W. Dean and A. Naibo, Recursive Functions, Fall 2024 Edition. Stanford University: Metaphysics Research Lab, Stanford University, 2024. Available: \href{https://plato.stanford.edu/archives/fall2024/entries/recursive-functions/}{https://plato.stanford.edu/archives/fall2024/entries/recursive-functions/}
        
            \bibitem{ChemTM} A. Hjelmfelt, E. D. Weinberger, and J. Ross, “Chemical implementation of neural networks and Turing machines.,” Proceedings of the National Academy of Sciences, vol. 88, no. 24, pp. 10983–10987, Dec. 1991, doi: https://doi.org/10.1073/pnas.88.24.10983.

            \bibitem{MCTM} “I Made a Working Computer with just Redstone!,” www.youtube.com. \href{https://www.youtube.com/watch?v=CW9N6kGbu2I}{https://www.youtube.com/watch?v=CW9N6kGbu2I}

            \bibitem{TerrariaTM} From Scratch, “I Made a 32-bit Computer Inside Terraria,” YouTube, Jun. 24, 2023. \href{https://www.youtube.com/watch?v=zXPiqk0-zDY}{https://www.youtube.com/watch?v=zXPiqk0-zDY}

            %This citation is weird because it wasn't being properly created from the ieee citation generator so i did it by hand. I hope it's okay
            \bibitem{TerrariaTMGH} misprit7, "computerraria", Feb. 16, 2023, GitHub repository, \href{https://github.com/misprit7/computerraria}{https://github.com/misprit7/computerraria}
            
            \bibitem{nand2tetris} “Home | nand2tetris,” nand2tetris, 2017. \href{https://www.nand2tetris.org/}{https://www.nand2tetris.org/}

            \bibitem{ELTCompSys} Noam Nisan, ELEMENTS OF COMPUTING SYSTEMS : building a modern computer from first principles. 2020.

            % The above sources are the actual journals and books read/used in the thesis.
            % the below sources are just citations for declarations of info which are not explicitly journals or formal publications

            \bibitem{WolframRiceThm} Sakharov, Alex. “Rice’s Theorem.” From MathWorld--A Wolfram Web Resource, created by Eric W. Weisstein, Wolfram.com, 2024. \href{https://mathworld.wolfram.com/RicesTheorem.html}{https://mathworld.wolfram.com/RicesTheorem.html}
            
            \bibitem{WikiRiceThm} Wikipedia Contributors, “Rice’s theorem,” Wikipedia, Sep. 24, 2019. \href{https://en.wikipedia.org/wiki/Rice%27s_theorem}{https://en.wikipedia.org/wiki/Rice\%27s\_theorem}
            
            \bibitem{BfWiki} “Brainfuck,” Wikipedia, Oct. 05, 2021. \href{https://en.wikipedia.org/wiki/Brainfuck}{https://en.wikipedia.org/wiki/Brainfuck}
            
            \bibitem{BfGH} 262588213843476, “Basics of BrainFuck,” Gist, Oct. 29, 2024. \href{https://gist.github.com/roachhd/dce54bec8ba55fb17d3a}{https://gist.github.com/roachhd/dce54bec8ba55fb17d3a}

            \bibitem{BfSO} speeder, “How does the Brainfuck Hello World actually work?,” Stack Overflow, May 30, 2013. \href{https://stackoverflow.com/questions/16836860/how-does-the-brainfuck-hello-world-actually-work/19869651#19869651}{https://stackoverflow.com/questions/16836860/how-does-the-brainfuck-hello-world-actually-work/19869651\#19869651}
                        
            \bibitem{SmallestBfCompiler} “brainfuck - Esolang,” Esolangs.org, 2023. \href{https://esolangs.org/wiki/Brainfuck#Self-interpreters}{https://esolangs.org/wiki/Brainfuck\#Self-interpreters}
            
            \bibitem{SmallestBfCompilerActual} Brainfuck.org, 2024. \href{https://brainfuck.org/dbfi.b}{https://brainfuck.org/dbfi.b}
            
            \bibitem{GitMadeinC} “git.git - The core git plumbing,” Kernel.org, 2024. \href{https://git.kernel.org/pub/scm/git/git.git/tree/}{https://git.kernel.org/pub/scm/git/git.git/tree/}

            \bibitem{LinuxMadeinC} “kernel/git/torvalds/linux.git - Linux kernel source tree,” Kernel.org, 2024. \href{https://git.kernel.org/pub/scm/linux/kernel/git/torvalds/linux.git/tree/}{https://git.kernel.org/pub/scm/linux/kernel/git/torvalds/linux.git/tree/}

            \bibitem{GNUPreferC}  “Source Language (GNU Coding Standards),” Gnu.org, 2024. \href{https://www.gnu.org/prep/standards/html_node/Source-Language.html}{https://www.gnu.org/prep/standards/html\_node/Source-Language.html}
            
            \bibitem{MarketShareOS} StatCounter, “Desktop Operating System Market Share Worldwide | StatCounter Global Stats,” StatCounter Global Stats, 2019. \href{https://gs.statcounter.com/os-market-share/desktop/worldwide/}{https://gs.statcounter.com/os-market-share/desktop/worldwide/}
            
            \bibitem{COBOLBanks} “On the past, present, and future of COBOL – Increment: Programming Languages,” increment.com. \href{https://increment.com/programming-languages/cobol-all-the-way-down/}{https://increment.com/programming-languages/cobol-all-the-way-down/}

            \bibitem{CSUNStudents} “How Does California State University--Northridge Rank Among America’s Best Colleges?,” @USNews, 2015. \href{https://www.usnews.com/best-colleges/california-state-university-northridge-1153}{https://www.usnews.com/best-colleges/california-state-university-northridge-1153}

            \bibitem{JSPopular} H. Shah, “7 Frontend JavaScript Frameworks Loved by Developers in 2022,” Simform - Product Engineering Company, Feb. 17, 2022. \href{https://www.simform.com/blog/javascript-frontend-frameworks/}{https://www.simform.com/blog/javascript-frontend-frameworks/}

            \bibitem{ChatGPTPython} D. K. K. has years of experience as a S. D. S. H. enjoys coding, teaching, and has created this website to make M. L. accessible to everyone, “Is ChatGPT Written In Python??? [We FINALLY Found The Proof]» EML,” enjoymachinelearning.com, Feb. 09, 2023. \href{https://enjoymachinelearning.com/blog/is-chatgpt-written-in-python/}{https://enjoymachinelearning.com/blog/is-chatgpt-written-in-python/}

            \bibitem{RDataSci} Simplilearn, “What is R: Overview, its Applications and what is R used for | Simplilearn,” Simplilearn.com, Oct. 25, 2021. \href{https://www.simplilearn.com/what-is-r-article}{https://www.simplilearn.com/what-is-r-article}

            \bibitem{DataSciLangs} J. C. Luna, “Learn R, Python \& Data Science Online,” www.datacamp.com, Mar. 2023. \href{https://www.datacamp.com/blog/top-programming-languages-for-data-scientists-in-2022}{https://www.datacamp.com/blog/top-programming-languages-for-data-scientists-in-2022}

            \bibitem{vonNeumannImg} Kapooht, “English: von Neumann Architecture,” Wikimedia Commons, Apr. 28, 2013. \href{https://commons.wikimedia.org/wiki/File:Von_Neumann_Architecture.svg}{https://commons.wikimedia.org/wiki/File:Von\_Neumann\_Architecture.svg}

            \bibitem{HarvardArchImg} “File:Harvard architecture.svg - Wikimedia Commons,” Wikimedia.org, May 11, 2010. \href{https://commons.wikimedia.org/wiki/File:Harvard_architecture.svg}{https://commons.wikimedia.org/wiki/File:Harvard\_architecture.svg}

            \bibitem{4080Specs} “NVIDIA GeForce RTX 4080 Specs,” TechPowerUp, Oct. 13, 2023. \href{https://www.techpowerup.com/gpu-specs/geforce-rtx-4080.c3888}{https://www.techpowerup.com/gpu-specs/geforce-rtx-4080.c3888}

            \bibitem{IntelSpecs} “Intel® CoreTM i9-13900KS Processor (36M Cache, up to 6.00 GHz) - Product Specifications,” Intel. \href{https://www.intel.com/content/www/us/en/products/sku/232167/intel-core-i913900ks-processor-36m-cache-up-to-6-00-ghz/specifications.html}{https://www.intel.com/content/www/us/en/products/sku/232167/intel-core-i913900ks-processor-36m-cache-up-to-6-00-ghz/specifications.html}

            \bibitem{ALUImg} “File:2-bit ALU.svg - Wikibooks, open books for an open world,” Wikibooks.org, Oct. 18, 2011. \href{https://en.wikibooks.org/wiki/File:2-bit_ALU.svg}{https://en.wikibooks.org/wiki/File:2-bit\_ALU.svg}

            \bibitem{CellAutWiki} Wikipedia Contributors, “Cellular automaton,” Wikipedia, Dec. 05, 2019. \href{https://en.wikipedia.org/wiki/Cellular_automaton}{https://en.wikipedia.org/wiki/Cellular\_automaton}

            \bibitem{CellAutWolfram} E. W. Weisstein, “Cellular Automaton,” From Mathworld--A Wolfram Web Resource, mathworld.wolfram.com. \href{https://mathworld.wolfram.com/CellularAutomaton.html}{https://mathworld.wolfram.com/CellularAutomaton.html}

            \bibitem{CellAutBook} A. Ilachinski, Cellular Automata. World Scientific, 2001.

            \bibitem{CGoLImg} “Math’s ‘Game of Life’ Reveals Long-Sought Repeating Patterns | Quanta Magazine,” Quanta Magazine, Jan. 18, 2024. \href{https://www.quantamagazine.org/maths-game-of-life-reveals-long-sought-repeating-patterns-20240118/}{https://www.quantamagazine.org/maths-game-of-life-reveals-long-sought-repeating-patterns-20240118/}

            \bibitem{Rule110Img} “File:One-d-cellular-automaton-rule-110.gif - Wikimedia Commons,” Wikimedia.org, Nov. 20, 2018. \href{https://commons.wikimedia.org/wiki/File:One-d-cellular-automaton-rule-110.gif}{https://commons.wikimedia.org/wiki/File:One-d-cellular-automaton-rule-110.gif}

            \bibitem{CalcTC} Code \& Optimism, “How to write a Turing-Complete Programming Language in 40 minutes in Ruby using Bable-Bridge,” YouTube, Sep. 19, 2012. \href{https://www.youtube.com/watch?v=_Uoyufkb5lk}{https://www.youtube.com/watch?v=\_Uoyufkb5lk}

            \bibitem{CBfInter} M. Kenyon, “How to Write a Brainfuck Interpreter in C\#,” Thesharperdev.com, Oct. 12, 2019. \href{https://thesharperdev.com/how-to-write-a-brainfuck-interpreter-in-c/}{https://thesharperdev.com/how-to-write-a-brainfuck-interpreter-in-c/}

            \bibitem{MeepWebsite} “Compiling to Brainf\#ck - Meep.,” InJuly, 2024. \href{https://injuly.in/blog/bfinbf/index.html}{https://injuly.in/blog/bfinbf/index.html}

            \bibitem{MeepGH} srijan-paul, “GitHub - srijan-paul/meep: A programming language that compiles to brainfuck.,” GitHub, 2020. \href{https://github.com/srijan-paul/meep}{https://github.com/srijan-paul/meep}

            \bibitem{PythonBfInt} M. Ueding, “Creating a Brainfuck interpreter,” Martin Ueding, Apr. 19, 2023. \href{https://martin-ueding.de/posts/creating-a-brainfuck-interpreter/}{https://martin-ueding.de/posts/creating-a-brainfuck-interpreter/}

            \bibitem{FuncProgrChap} “Functional Programming,” learn.saylor.org. \href{https://learn.saylor.org/mod/book/tool/print/index.php?id=33044&chapterid=13087}{https://learn.saylor.org/mod/book/tool/print/\-index.php?id=33044\&chapterid=13087}

            \bibitem{LambdaCalcWiki} Wikipedia Contributors, “Lambda calculus,” Wikipedia, Jan. 02, 2020. \href{https://en.wikipedia.org/wiki/Lambda_calculus}{https://en.wikipedia.org/wiki/Lambda\_calculus}

            \bibitem{LambdaFuncsList} “Collected Lambdas,” jwodder.freeshell.org. \href{https://jwodder.freeshell.org/lambda.html}{https://jwodder.freeshell.org/lambda.html}

            \bibitem{SKICalcWiki} Wikipedia Contributors, “SKI combinator calculus,” Wikipedia, Oct. 28, 2024. \href{https://en.wikipedia.org/wiki/SKI_combinator_calculus}{https://en.wikipedia.org/wiki/SKI\_combinator\_calculus}

            \bibitem{RedditLambdaCalcPost} “Reddit - Dive into anything,” Reddit.com, 2017. \href{https://www.reddit.com/r/ProgrammerHumor/comments/78z90f/when_you_need_to_know_if_a_number_is_even_or_odd/}{https://www.reddit.com/r/ProgrammerHu\-mor/comments/78z90f/when\_you\_need\_to\_know\_if\_a\_number\_is\_even\_or\_odd/} 

            \bibitem{RedditLambdaCalcComment} “Reddit - Dive into anything,” Reddit.com, 2017. \href{https://www.reddit.com/r/ProgrammerHumor/comments/78z90f/comment/doylzry/}{https://www.reddit.com/r/ProgrammerHu\-mor/comments/78z90f/comment/doylzry/}

            \bibitem{TruncJava} “What is truncation in Java - Javatpoint,” www.javatpoint.com, 2021. \href{https://www.javatpoint.com/what-is-truncation-in-java}{https://www.javatpoint.com/what-is-truncation-in-java}

            \bibitem{TruncC} GeeksforGeeks, “trunc() , truncf() , truncl() in C language,” GeeksforGeeks, Jan. 25, 2018. \href{https://www.geeksforgeeks.org/trunc-truncf-truncl-c-language/}{https://www.geeksforgeeks.org/trunc-truncf-truncl-c-language/}

            \bibitem{ExcelTC} “The Excel Formula Language Is Now Turing-Complete,” InfoQ. \href{https://www.infoq.com/articles/excel-lambda-turing-complete/}{https://www.infoq.com/articles/excel-lambda-turing-complete/}

            \bibitem{AccTC} “Accidentally Turing-Complete,” beza1e1.tuxen.de. \href{https://beza1e1.tuxen.de/articles/accidentally_turing_complete.html}{https://beza1e1.tuxen.de/articles/accident\-ally\_turing\_complete.html}

            \bibitem{AccTC2} G. Branwen, “Surprisingly Turing-Complete,” gwern.net, Dec. 2012, Available: \href{https://gwern.net/turing-complete}{https://gwern.net/turing-complete}

            \bibitem{SecVuln} “A deep dive into an NSO zero-click iMessage exploit: Remote Code Execution,” Blogspot.com, 2021. \href{https://googleprojectzero.blogspot.com/2021/12/a-deep-dive-into-nso-zero-click.html?m=1}{https://googleprojectzero.blogspot.com/2021/12/a-deep-dive-into-nso-zero-click.html?m=1}

            \bibitem{ActorModelParallel} Wikipedia Contributors, “Actor model,” Wikipedia, Nov. 14, 2019. \href{https://en.wikipedia.org/wiki/Actor_model}{https://en.wikipedia.org/wiki/Actor\_model}
           
            \bibitem{ActorJavaParallel} L. Nigro, “Parallel Theatre: An actor framework in Java for high performance computing,” Simulation Modelling Practice and Theory, vol. 106, p. 102189, Jan. 2021, doi: https://doi.org/10.1016/j.simpat.2020.102189. 

            \bibitem{CartesianPlane} “Cartesian Plane - Definition, Meaning, Quadrants, Examples,” Cuemath. https://www.cuemath.com/geometry/cartesian-plane/

            \bibitem{JavaUse1} “What Is Java Used For: 12 Real World Java Applications,” www.softwaretestinghelp.com. https://www.softwaretestinghelp.com/real-world-applications-of-java/

            \bibitem{JavaUse2} “What is Java Used For?,” Codecademy News, Jun. 24, 2021. https://www.codecademy.com/resources/blog/what-is-java-used-for/

            \bibitem{CPPUse1} Simplilearn, “Top 7 Practical Applications of C++ and the Way to Build a Career in the Field,” Simplilearn.com, Jan. 16, 2020. https://www.simplilearn.com/c-plus-plus-programming-for-beginners-article

            \bibitem{CPPUse2} “Uses of C++ | 10 Reasons Why You Should Use C++,” EDUCBA, Jul. 19, 2019. https://www.educba.com/uses-of-c-plus-plus/

            \bibitem{CPPFirefox} “mozilla/gecko-dev,” GitHub, Feb. 01, 2024. https://github.com/mozilla/gecko-dev

            \bibitem{FirefoxMain} “The new, fast browser for Mac, PC and Linux | Firefox,” Mozilla, 2000. https://www.mozilla.org/en-US/firefox/
        
            \bibitem{PHPScriptingServer} Astari S., “What Is PHP? Learning All About the Scripting Language,” Hostinger Tutorials, Apr. 30, 2019. https://www.hostinger.com/tutorials/what-is-php/

            \bibitem{PHPWP} http://facebook.com/syedbalkhi, “What is PHP? How PHP is Used in WordPress?,” WPBeginner, 2019. https://www.wpbeginner.com/glossary/php/

            \bibitem{PHPLaravel} T. Otwell, “Laravel - The PHP Framework For Web Artisans,” Laravel.com, 2015. https://laravel.com/

            \bibitem{TopLangs1} L. dev, “Top 8 Most Demanded Programming Languages in 2023,” Devjobsscanner, Jun. 22, 2023. https://www.devjobsscanner.com/blog/top-8-most-demanded-programming-languages/

            \bibitem{TopLangs2} L. Whitney, “Top 10 programming languages employers want in 2023,” TechRepublic, Feb. 03, 2023. https://www.techrepublic.com/article/top-programming-languages-employers-want/

            \bibitem{TopLangs3} “11 of the Most In-Demand Coding Languages,” Indeed Career Guide. https://www.indeed.com/career-advice/career-development/most-in-demand-coding-languages

            \bibitem{PyLibs} “Search results,” PyPI. https://pypi.org/search/

            \bibitem{NPM} npm, “npm | build amazing things,” Npmjs.com, 2019. https://www.npmjs.com/

            \bibitem{Yarn} Yarnpkg.com, 2019. https://yarnpkg.com/

            \bibitem{EasyLangs1} “10 Hardest and Easiest Programming Languages in 2024 - GUVI Blogs,” Mar. 01, 2023. https://www.guvi.in/blog/easiest-programming-languages-to-hardest-ranked/

            \bibitem{EasyLangs2} “The 10 Most Popular Coding Languages to Learn in 2023 | BestColleges,” www.bestcolleges.com. https://www.bestcolleges.com/bootcamps/guides/most-important-coding-languages/

            \bibitem{FP1} J. M. Fernandes, “Functional Programming: With Examples and Lots of Cats,” Medium, Sep. 29, 2021. https://medium.com/arctouch/thinking-functional-now-with-example-and-cats-8b9c2478b9af            

            \bibitem{FP2} GeeksForGeeks, “Functional Programming Paradigm - GeeksforGeeks,” GeeksforGeeks, Jan. 02, 2019. https://www.geeksforgeeks.org/functional-programming-paradigm/

            \bibitem{RefTransp1} “Referential transparency,” Wikipedia, Jan. 05, 2021. https://en.wikipedia.org/wiki/Referential\_transparency

            \bibitem{RefTransp2} Claudiu, “What is referential transparency?,” Stack Overflow, Oct. 17, 2008. https://stackoverflow.com/a/9859966

            \bibitem{ElixirList} beam-community, “elixir-companies/priv/companies at main · beam-community/elixir-companies,” GitHub, 2015. https://github.com/beam-community/elixir-companies/tree/main/priv/companies

            \bibitem{HaskellListReadme} erkmos, “GitHub - erkmos/haskell-companies: A gently curated list of companies using Haskell in industry,” GitHub, 2017. https://github.com/erkmos/haskell-companies?tab=readme-ov-file

            \bibitem{ElixirListReadme} beam-community, “GitHub - beam-community/elixir-companies: A list of companies currently using Elixir in production.,” GitHub, 2015. https://github.com/beam-community/elixir-companies

            \bibitem{HaskellList} “Haskell in industry - HaskellWiki,” Haskell.org, 2018. https://wiki.haskell.org/Haskell\_in\_industry

            \bibitem{OCamlCoq} coq, “GitHub - coq/coq: Coq is a formal proof management system. It provides a formal language to write mathematical definitions, executable algorithms and theorems together with an environment for semi-interactive development of machine-checked proofs.,” GitHub, Sep. 04, 2024. https://github.com/coq/coq?tab=readme-ov-file

            \bibitem{ErlangUses} F. Cesarini, “Companies Who Use Erlang,” Erlang Solutions, Sep. 11, 2019. https://www.erlang-solutions.com/blog/which-companies-are-using-erlang-and-why-mytopdogstatus/

        \end{thebibliography}

    \end{singlespace}

    \appendix
    \addcontentsline{toc}{chapter}{Appendix}
        
    \input{ProofAssistants}
    
    \input{CodeSegments}

\end{document}